---
title: "Statistical Analysis of Developer's Survey"
output:
  word_document:
  html_document:default
    df_print: paged
---

Installing Required Packages
```{r}
required_packages <- c('tidyverse', 'ggplot2', 'mosaic', 'kableExtra', 'ggpmisc', 'bootstrap')
for (p in required_packages) {
  if(!require(p,character.only = TRUE)) {
    install.packages(p, dep = TRUE)
  }
}

```

Reading Data
```{r}
developer_survey<- read_csv("/Users/shuaibismail/Documents copy/803/final project/survey_results_public.csv", show_col_types = FALSE)

head(developer_survey)
```


Exploratory Data Analysis
```{r}
developer_survey<- as_tibble(developer_survey)
```


```{r}
developer_survey<- developer_survey%>%
  filter(Country == "United States")
```


```{r}
xtabs(~Gender,data=developer_survey)
```
```{r}
prop.table(table(developer_survey$Gender))
```


```{r}
prop.table(table(developer_survey$UndergradMajor))
```


```{r}
prop.table(table(developer_survey$FormalEducation))
```


```{r}
prop.table(table(developer_survey$YearsCodingProf))
```


```{r}
YearsCodingProf<- developer_survey$YearsCodingProf%>%
  mutate(CodingExperience = recode_factor(YearsCodingProf, `1` = "0-2 years",`2` = "3-5 years", `3` ="6-8 years", `4` = "9-11 years", `5` = "12-14 years", `6` = "15-17 years", `7` = "18-20 years", `8` = "21-23 years", `9` = "24-26 years", `3` = "27-29 years", `5` = "30 or more years"))
```


```{r}
table(developer_survey$ConvertedSalary, developer_survey$YearsCodingProf)
```


```{r}
table(developer_survey$UndergradMajor, developer_survey$Gender)
```


```{r}
summary(developer_survey)
```


```{r}
favstats(~ConvertedSalary,data=developer_survey)
```
Population Mean for Salary = 154413.1	and median is 100,000

```{r}
summary(developer_survey$FormalEducation)
```

Proportionate Stratified Sampling
```{r}
set.seed(123)
Masters_samp<- developer_survey%>%
  filter(FormalEducation=="Master’s degree (MA, MS, M.Eng., MBA, etc.)")%>%
  select(FormalEducation)%>%
  sample_frac(0.4)
  
Masters_samp
```


```{r}
set.seed(123)
Bachelor_samp<- developer_survey%>%
  filter(FormalEducation=="Bachelor’s degree (BA, BS, B.Eng., etc.)")%>%
  select(FormalEducation)%>%
  sample_frac(0.6)

Bachelor_samp
```


```{r}
set.seed(123)
Secondary_samp<- developer_survey%>%
  filter(FormalEducation=="Secondary school (e.g. American high school, German Realschule or Gymnasium, etc.)")%>%
  select(FormalEducation)%>%
  sample_frac(0.2)

Associate_samp<- developer_survey%>%
  filter(FormalEducation=="Associate degree")%>%
  select(FormalEducation)%>%
  sample_frac(0.2)


doctoral_samp<- developer_survey%>%
  filter(FormalEducation=="Other doctoral degree (Ph.D, Ed.D., etc.)")%>%
  select(FormalEducation)%>%
  sample_frac(0.1)


Some_college_samp<- developer_survey%>%
  filter(FormalEducation=="Some college/university study without earning a degree")%>%
  select(FormalEducation)%>%
  sample_frac(0.2)


Primary_samp<- developer_survey%>%
  filter(FormalEducation=="Primary/elementary school")%>%
  select(FormalEducation)%>%
  sample_frac(0.1)


Professional_samp<- developer_survey%>%
  filter(FormalEducation=="Professional degree (JD, MD, etc.)")%>%
  select(FormalEducation)%>%
  sample_frac(0.1)


No_education_samp<- developer_survey%>%
  filter(FormalEducation=="I never completed any formal education")%>%
  select(FormalEducation)%>%
  sample_frac(0.1)

Education_samp<- rbind(No_education_samp, Professional_samp, Primary_samp, Some_college_samp,doctoral_samp, Associate_samp, Secondary_samp, Bachelor_samp, Masters_samp)
                                       
Education_samp                                
                            
```


Salary Sampling
```{r}
set.seed(123)
MastersSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Master’s degree (MA, MS, M.Eng., MBA, etc.)")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.4)

BacheloSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Bachelor’s degree (BA, BS, B.Eng., etc.)")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.4)

set.seed(123)
SecondarySalary_samp<- developer_survey%>%
  filter(FormalEducation=="Secondary school (e.g. American high school, German Realschule or Gymnasium, etc.)")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.2)

AssociateSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Associate degree")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.2)


doctoralSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Other doctoral degree (Ph.D, Ed.D., etc.)")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.1)


Some_collegeSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Some college/university study without earning a degree")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.2)


PrimarySalary_samp<- developer_survey%>%
  filter(FormalEducation=="Primary/elementary school")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.1)


ProfessionalSalary_samp<- developer_survey%>%
  filter(FormalEducation=="Professional degree (JD, MD, etc.)")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.1)


No_educationSalary_samp<- developer_survey%>%
  filter(FormalEducation=="I never completed any formal education")%>%
  select(ConvertedSalary)%>%
  sample_frac(0.1)

Salary_samp<- rbind(No_educationSalary_samp, ProfessionalSalary_samp, PrimarySalary_samp, Some_collegeSalary_samp, doctoralSalary_samp, AssociateSalary_samp, SecondarySalary_samp, BacheloSalary_samp, MastersSalary_samp)

Salary_samp

```


Descriptive Analysis
```{r}
qqnorm(Salary_samp, frame = FALSE)
qqline(Salary_samp, col = "steelblue", lwd = 2)
```


```{r}
hist(Salary_samp, xlab="Salary Range", main="Histogram for Salaries of Developers", col="blue", prob=TRUE, breaks=100)
xfit<-seq(min(s),max(s))
yfit<-dnorm(xfit,mean(s), sd(s))
lines(xfit,yfit,col="red", lwd=3)
```


Inferential Analysis
```{r}
## Methods of Moments
n=6689
MM.mu=sum(Salary_samp)/n
MM.mu

sigma2.mu=sum((Salary_samp-MM.mu)^2)/n
sigma2.mu 

## Maxmimum Likelihood Estimation
MLE.mu=sum(Salary_samp)/n
MLE.mu

sigmaMLE=sum((Salary_samp-MM.mu)^2)/n
sigmaMLE

#For Normal Distribution Method of Moment and Maximum Likelihood Estimators are equal.
```

Jackknife Method
```{r}
as.data.frame(Salary_samp)
data=Salary_samp
n=6689
mean.data=mean(data)
jk=sapply(1:n, 
          function(i) mean(data[-i])
)
mean.jk=mean(jk)
se.jk=sqrt((n-1)*mean((jk-mean.data)^2))
list(mean=mean.jk, se=se.jk)
```


95% Confidence Intervals
```{r}
Salary_samp
Norm_CI=function(x, alpha){
n=6689
df=n-1
ME=qt(1-alpha/2, df)*sd(x)/sqrt(n)
c(Lower=mean(x)-ME, Upper=mean(x)+ME)
}

Norm_CI(Salary_samp, 0.05)

```
```{r}

Sigma.CI=function(x,alpha){
n=length(x)
df=n-1
s2=sd(x)^2
c(Lower=df*s2/qchisq(1-alpha/2, df), Upper=df*s2/qchisq(alpha/2, df))
}

Sigma.CI(Salary_samp, 0.05)
```

Hypothesis Test
```{r}
test.stat=(mean(Salary_samp)-household_income)/(sqrt(var(Salary_samp)/length(Salary_samp)))
pval=pnorm(test.stat)
list(test.stat=test.stat, pval=pval)
```
High P-value means little or no emperical evidence against the null hypothesis

1.Low pvalue: strong empirical evidence against h0
2.High pvalue: little or 'no' empirical evidence against h0

Right Tail Test
```{r}
#Right Tail Test
t.test.right <- function(data, mu0, alpha){
    t.stat <- (mean(data) - mu0) / (sqrt(var(data) / length(data)))
    dof <- length(data) - 1
    t.critical <- qt(1-alpha, df= dof) #Es alpha 0.05 -> 1.64 (df=Inf)
    p.value <- 1 - pt(t.stat, df= dof)
    if(t.stat >= t.critical)
    {
        print("Reject H0")
    }
    else
    {
        print("Accept H0")
    }
    print('T statistic')
    print(t.stat)
    print('T critical value')
    print(t.critical)
    print('P value')
    print(p.value)
    print("#####################")
    return(t.stat)
}
    t.test.right(Salary_samp, household_income, 0.05)
```

Regression Analysis
```{r}
d=bind_cols(Education_samp, Salary_samp=unlist(Salary_samp))

Education_samp2=sample(Education_samp, 6689, replace = FALSE)
YearsCodingProf2=sample(developer_survey$YearsCodingProf, 6689, replace = FALSE)

Salary_samp=unlist(Salary_samp)

d <- data.frame(Education_samp2,Salary_samp)

Education_samp=unlist(Education_samp)

model=lm(Salary_samp~Education_samp2, data=d)
model
```


```{r}
summary(model)
```


```{r}
 mlr=lm(Salary_samp~Education_samp2+YearsCodingProf2)
summary(mlr)
```


```{r}
qqplot(mlr)
```


```{r}
household_income=63179

set.seed(123)
data=Salary_samp
test_stat=(mean(Salary_samp)-household_income)/sd(Salary_samp)/sqrt(length(Salary_samp))
pval=pnorm(test_stat)
list(test_stat=test.stat, pval=pval)
```


```{r}
data=stratified_sample$HINCP
test.stat=(mean(data)-80000)/(sd(data)/sqrt(length(stratified_sample$HINCP)))

1-pt(test.stat,length(stratified_sample$HINCP-1))
```




```{r}
library(bootstrap)
temp <- jackknife(data, mean)
c(temp$jack.bias, temp$jack.se)
```


```{r}
data=Salary_samp-1
B=10000
n=nrow(Salary_samp)
resample.boot=matrix(sample(data, size = B*n, replace=TRUE), B, n)
mean.boot=apply(resample.boot, 1, mean)
se=sd(mean.boot)
se

```


```{r}
data=Education_samp

ggplot(data.frame(Education_samp), aes(x=Bachelor_samp))+
  geom_histogram(binwidth=25, aes(y=..density..))+
  geom_density(color="red")
```


```{r}
mean.b = function(K, Salary_samp){
  resamples = lapply(1:K, function(i) sample(data, replace=T))
  mean.r = sapply(resamples, mean)
  se.b= sd(mean.r)
  list(se=se.b, means=mean.r, mean=mean(mean.r))
                     
}
```

Conclusion

Average developer Income is greater than $63136 for selected sample population, as proved using MLE, Confidence Interval and Hypothesis Testing 

Both our simple linear model and multi-linear model show that our independent variable, education level can explain part of our dependent variable, developer income. Yet, neither simple linear nor the multi-linear model is an ideal fit for our sample data. Therefore designed model does not fit developer Income and Education Attainment. 

Our research concerning relationship between education level and developer income is yet waiting for better model fit and more authoritative explanation. 




